Standartox: Standardizing toxicity data

### 2020-04-10 article/article.tex
Updates from Overleaf

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz
%% Additional packages
\usepackage{standalone}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{rotating}
\usepackage{tikzpeople}
\usepackage{makecell}
\usepackage{refcheck} % debuging
% code
\lstset{
    language = R,
    basicstyle = \small\ttfamily,
    stringstyle = \color{orange},
    % key words
    otherkeywords = {0,1,2,3,4,5,6,7,8,9},
    morekeywords = {TRUE,FALSE},
    keywordstyle = \color{blue},
    % comments
    commentstyle = \color{darkgray},
    morecomment = [f][\color{red}][0]{\#},
    % frame
    frame = single,
    backgroundcolor = \color{white}
}
% tikz
\usetikzlibrary{trees,
                arrows,
                shapes.geometric,
                positioning,
                calc,
                backgrounds,
                fit}

@@

\subsection{Accuracy Assessment}
To validate Standartox results we compared geometric means resulting from the aggregation in Standartox to the corresponding values from other databases, for chemicals where data was available in both resources. The PPDB provides ecotoxicity data on a few selected species from chemical risk assessment that have been manually quality controlled through expert judgement \citep{lewis_international_2016}. The vast majority of aggregated values (91.9~\%) of Standartox lie within one order of magnitude of the corresponding PPDB values (n~=~3601). This would increase to 92.6~\%, when restricting the comparison to Standartox values where data from at least five experiments are available. Similarly, we compared Standartox to ecotoxicity values for \textit{D. magna} from the ChemProp \citep{ufzdepartmentofecologicalchemistry_chemprop_2019} software, which estimates LC\textsubscript{50} values via quantitative structure-activity relationship (QSAR) models \citep{schuurmann_quantitative_2011}. We found that 95~\% of Standartox values lie within one order of magnitude of the ChemProp (n~=~179) values. However, the diﬀerence is not necessarily an indication of lower quality of Standartox estimates but may also reflect the wider range of experimental conditions for which data are available in the database underlying Standartox as well as inaccurate predictions for QSAR models, respectively (Figure~\ref{fig:standartox_ppdb_diff}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{The authors thank Eduard Szöcs for inspiration through a blog post on ho to build a local EPA Ecotox database.}

@@

%=====================================
% References, variant B: external bibliography
%=====================================
\reftitle{References}
\nocite{*} % debuging
\externalbibliography{yes}
\bibliography{refs/references-standartox.bib,
              refs/references-standartox-rpackages.bib}

>>>>

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz
%% Additional packages
\usepackage{standalone}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{rotating}
\usepackage{tikzpeople}
\usepackage{makecell}
% \usepackage{refcheck} % debuging
% code
\lstset{
    language = R,
    basicstyle = \small\ttfamily,
    stringstyle = \color{orange},
    % key words
    otherkeywords = {0,1,2,3,4,5,6,7,8,9},
    morekeywords = {TRUE,FALSE},
    keywordstyle = \color{blue},
    % comments
    commentstyle = \color{darkgray},
    morecomment = [f][\color{red}][0]{\#},
    % frame
    frame = single,
    backgroundcolor = \color{white}
}
% tikz
\usetikzlibrary{trees,
                arrows,
                shapes.geometric,
                positioning,
                calc,
                backgrounds,
                fit}

@@

\subsection{Accuracy Assessment}
To validate Standartox results we compared geometric means resulting from the aggregation in Standartox to the corresponding values from other databases, for chemicals where data was available in both resources. The PPDB provides ecotoxicity data on a few selected species from chemical risk assessment that have been manually quality controlled through expert judgement \citep{lewis_international_2016}. The vast majority of aggregated values (91.9~\%) of Standartox lie within one order of magnitude of the corresponding PPDB values (n~=~3601). This would increase to 92.6~\%, when restricting the comparison to Standartox values where data from at least five experiments are available. Similarly, we compared Standartox to ecotoxicity values for \textit{D. magna} from the ChemProp \citep{ufzdepartmentofecologicalchemistry_chemprop_2019} software, which estimates LC\textsubscript{50} values via quantitative structure-activity relationship (QSAR) models \citep{schuurmann_quantitative_2011}. We found that 95~\% of Standartox values lie within one order of magnitude of the ChemProp (n~=~179) values. However, the difference is not necessarily an indication of lower quality of Standartox estimates but may also reflect the wider range of experimental conditions for which data are available in the database underlying Standartox as well as inaccurate predictions for QSAR models, respectively (Figure~\ref{fig:standartox_ppdb_diff}).

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{The authors thank Eduard Szöcs for inspiration through a blog post on how to build a local version of the EPA Ecotox database.}

@@

%=====================================
% References, variant B: external bibliography
%=====================================
\reftitle{References}
% \nocite{*} % debuging
\externalbibliography{yes}
\bibliography{refs/references-standartox.bib,
              refs/references-standartox-rpackages.bib}
